---
title: "figures_for_popgroup"
output: html_document
date: "2024-12-06"
---

Other people running this will need to adapt all the paths to various files (all available on github), shouldn't be too bad though

```{r}
library(tidyverse)
library(gridExtra)
library(grid)
```



```{r}
raw_genome_1 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_1_genome_info.tsv",delim="\t") %>% mutate(reads_source = 1)
raw_genome_3 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_3_genome_info.tsv",delim="\t") %>% mutate(reads_source = 3)
raw_genome_7 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_7_genome_info.tsv",delim="\t") %>% mutate(reads_source = 7)
raw_genome_10 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_10_genome_info.tsv",delim="\t") %>% mutate(reads_source = 10)
raw_genome_25 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_25_genome_info.tsv",delim="\t") %>% mutate(reads_source = 25)
raw_genome_M5_3 <- rbind(raw_genome_1, raw_genome_3, raw_genome_7, raw_genome_10, raw_genome_25)
```



lifestyle plus phylo joins the MAG name with some limited taxonomy and also an "eye-balled" categorization of what lifestyle (surface, generalist, subsurface) that MAG subscribes to 
```{r}
lifestyle_plus_phylo <- read_delim("~/Projects/Sediments/lifestyles.tsv",delim="\t")
```
```{r}
raw_genome_M5_3 <- left_join(raw_genome_M5_3 %>% mutate(genome = gsub(".filtered.fa","", gsub("refs/","", genome))), lifestyle_plus_phylo)
```


```{r}
raw_genome_M5_3 %>% ggplot(aes(x=factor(reads_source), y=coverage, color=genome)) + geom_point(size=2.5) + theme_bw() + theme(text=element_text(size=22)) + labs(x="Reads Source (CmBSF)", color="MAG")
```
```{r}
#raw_genome_M5_3 %>% ggplot(aes(x=factor(reads_source), y=coverage, color=genome)) + geom_point() + theme_bw() + theme(text=element_text(size=22)) + labs(x="Reads Source (CmBSF)", color="MAG")
```

proc genes -> filtered plus normalized
```{r}
proc_genes_1 <- read_delim("~/Projects/Sediments/all_depths/IS_proc_M5_3_1_gene_info.tsv",delim="\t")%>% mutate(reads_source = 1)
proc_genes_3 <- read_delim("~/Projects/Sediments/all_depths/IS_proc_M5_3_3_gene_info.tsv",delim="\t")%>% mutate(reads_source = 3)
proc_genes_7 <- read_delim("~/Projects/Sediments/all_depths/IS_proc_M5_3_7_gene_info.tsv",delim="\t")%>% mutate(reads_source = 7)
proc_genes_10 <- read_delim("~/Projects/Sediments/all_depths/IS_proc_M5_3_7_gene_info.tsv",delim="\t")%>% mutate(reads_source = 10)
proc_genes_25 <- read_delim("~/Projects/Sediments/all_depths/IS_proc_M5_3_7_gene_info.tsv",delim="\t")%>% mutate(reads_source = 25)

proc_genes <- rbind(proc_genes_1, proc_genes_3, proc_genes_7, proc_genes_10, proc_genes_25)
```


This next plot


```{r}
genes_3_v_25 <- proc_genes %>%
  filter((reads_source == 3) | (reads_source == 25)) %>%
  select(scaffold, gene,reads_source, coverage, pNpS_variants) %>% 
  pivot_wider(names_from = reads_source, values_from = c(coverage, pNpS_variants))
```

```{r}
genes_3_v_25 %>% ggplot(aes(x=log10(coverage_3+0.01), y=log10(coverage_25+0.01)))+
  geom_hex() +
  geom_hline(yintercept = log10(5), color="red",linetype="dotted",linewidth=1) +
  geom_vline(xintercept = log10(5),color="red",linetype="dotted",linewidth=1 ) + 
  scale_fill_continuous(high="darkblue",low="lightblue") + 
  theme_bw()
```

```{r}
stb <- read_delim("~/Projects/Sediments/all_depths/combined_reference_M5_3.stb",delim="\t",col_names = c("scaffold","genome")) %>% mutate(genome = gsub(".filtered.fa","", genome))
head(stb)
genes_3_v_25 <- left_join(genes_3_v_25, stb)
```
```{r}
genes_3_v_25 <- left_join(genes_3_v_25,lifestyle_plus_phylo)
```

```{r}
attempt_1 <- genes_3_v_25 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
 # group_by(lifestyle) %>% 
#  reframe(
  ggplot(data=., aes(x=pNpS_variants_25, y=pNpS_variants_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(aes(
      high=ifelse(
        lifestyle == "Surface Specialist",
        "#E97132",
        ifelse(
          lifestyle == "Generalist",
          "#595959",
          "#156082"
        )
      ),
      low=ifelse(
        lifestyle=="Surface specialist",
       "#f8cfba",
       ifelse(
         lifestyle =="Generalist",
          "#d9d9d9",
          "#bde3f4")
       ))) +
    facet_wrap(~lifestyle)
 # ) %>% .$subplot %>% arrangeGrob(grobs = ., nrow = 1) %>% grid.arrange()
```


This plot uses IS's pN/pS rather than theta-hat-N theta-hat-S

```{r}
surface_subplot <- genes_3_v_25 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Surface Specialist") %>%
  ggplot(aes(x=pNpS_variants_25, y=pNpS_variants_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
    coord_fixed() + 
    geom_abline(slope=1, intercept=0, color="red",linewidth=1) +
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#E97132",
      low= "#f8cfba",
      ) +
    theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        text=element_text(size=22)) +
  labs(title="Surface Specialists")

# +
  #  facet_wrap(~lifestyle)
generalist_subplot <- genes_3_v_25 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Generalist") %>%
  ggplot(aes(x=pNpS_variants_25, y=pNpS_variants_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
    coord_fixed() + 
  geom_abline(slope=1, intercept=0, color="red",linewidth=1) +
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#595959",
      low= "#d9d9d9",
      ) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        text=element_text(size=22)) +
  labs(title="Generalists")
subsurface_subplot <- genes_3_v_25 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Subsurface Specialist") %>%
  ggplot(aes(x=pNpS_variants_25, y=pNpS_variants_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
  geom_abline(slope=1, intercept=0, color="red",linewidth=1) +
  coord_fixed() + 
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#156082",
      low= "#bde3f4",
      ) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        text=element_text(size=22)) +
  labs(title="Subsurface Specialists")

grid.arrange(surface_subplot, generalist_subplot, subsurface_subplot,nrow=2, bottom=textGrob("pN/pS 25cm", gp=gpar(fontsize=22)),left=textGrob("pN/pS 3cm",gp=gpar(fontsize=22),rot=90))
```
```{r}
raw_genome_M5_3 %>% ggplot(aes(x=genome, y=factor(reads_source), fill=log10(coverage))) + geom_tile(width=0.9) + scale_fill_continuous(high="#79206f",low="#f4d7f1")  + scale_x_discrete(guide = guide_axis(angle = 60)) + theme(axis.title = element_text(size=22), legend.title = element_text(size=22),axis.text.x = element_blank(), axis.text.y=element_text(size=20)) + labs(x="MAGs", y="Reads Source (CmBSF)") + scale_y_discrete(limits=rev)
```

```{r}
rg_w_theta <- read_delim("~/Projects/Sediments/raw_genes_filtered_w_theta_hat.tsv",delim="\t")
head(rg_w_theta)
```
```{r}
rg_w_theta <- rg_w_theta %>% mutate(theta_ns = theta_n / theta_s)
```

```{r}
rg_w_theta_325 <- rg_w_theta %>% filter((reads_source == 3) | (reads_source == 25)) %>% select(MAG, gene,  scaffold, reads_source, coverage, theta_ns) %>% pivot_wider(names_from = reads_source, values_from = c(theta_ns, coverage))
rg_w_theta_325 <- left_join(rg_w_theta_325 %>% mutate(genome = gsub(".filtered.fa","", gsub("refs/","", MAG))), lifestyle_plus_phylo)
```


This plot here is theta-hat-N / theta-hat-S by depth 

```{r}
surface_subplot_theta <- rg_w_theta_325 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Surface Specialist") %>%
  ggplot(aes(x=theta_ns_25, y=theta_ns_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
    coord_fixed() + 
    geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#e76118",
      low= "#f8cfba",
      ) +
    theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
       # legend.position="none",
        text=element_text(size=22)) +
  labs(title="Surface Specialists")

# +
  #  facet_wrap(~lifestyle)
generalist_subplot_theta <- rg_w_theta_325 %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Generalist") %>%
  ggplot(aes(x=theta_ns_25, y=theta_ns_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
    coord_fixed() + 
  geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#494949",
      low= "#d9d9d9",
      ) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        #legend.position="none",
        text=element_text(size=22)) +
  labs(title="Generalists")
subsurface_subplot_theta <- rg_w_theta_325  %>% filter((coverage_3 > 5) & (coverage_25 > 5)) %>% 
  filter(lifestyle =="Subsurface Specialist") %>%
  ggplot(aes(x=theta_ns_25, y=theta_ns_3)) +
    geom_hex() +
    xlim(0,1) +
    ylim(0,1) +
    theme_bw() + 
  geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
  coord_fixed() + 
    guides(fill = guide_colourbar(title.position = "top")) +
    scale_fill_continuous(
      high="#12516e",
      low= "#bde3f4",
      ) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
       # legend.position="none",
        text=element_text(size=22)) +
  labs(title="Subsurface Specialists")

grid.arrange(surface_subplot_theta, generalist_subplot_theta, subsurface_subplot_theta,nrow=2, bottom=textGrob("\U03B8N / \U03B8S 25cm", gp=gpar(fontsize=22)),left=textGrob("\U03B8N / \U03B8S 3cm",gp=gpar(fontsize=22),rot=90))
```

This next plot is theta-hat-N/theta-hat-S for all depths.  The hypothesis here is that maybe 7 & 10cmbsf have relaxed selection/positive selection as they transition from the oxygenated surface sediments to the anoxic/micro-oxic subsurface regime


```{r}
rg_w_theta <- left_join(rg_w_theta %>% mutate(genome = gsub(".filtered.fa","", gsub("refs/","", MAG))), lifestyle_plus_phylo)

rg_w_theta %>% filter(coverage > 5) %>% ggplot(aes(x=genome, y=log10(theta_ns+0.01), color=factor(reads_source))) + geom_boxplot() + facet_wrap(~lifestyle,scales="free_x") + theme_bw() + labs(y="log10 \U03B8N / \U03B8S + 0.01", color="CmBSF")

```


This section creates rg_w_theta_sum (rg -> raw genes, sum -> summary, that is by MAG) and the same but in even/odd codon form (even/odd -> eo).  The rest of these plots are pretty self explanatory for a bit

```{r}

rg_w_theta_sum <- rg_w_theta %>% group_by(MAG, reads_source, lifestyle) %>% filter(coverage > 5) %>% filter(theta_ns != Inf) %>% summarise(
  theta_s_mean = mean(theta_s, na.rm=T),
  theta_s_median = median(theta_s, na.rm=T),
  theta_ns_mean = mean(theta_ns, na.rm=T),
  theta_ns_median = median(theta_ns, na.rm=T),
  n = n()
)
rg_w_theta_sum_eo <- left_join(
  rg_w_theta %>% group_by(MAG, reads_source, lifestyle) %>% filter(coverage > 5) %>% filter(theta_ns != Inf) %>%
  filter((start %% 2 ) == 0) %>% #even numbered starts -> theta-hat S, odd numbered starts -> theta-hat NS
  summarise(
    theta_s_mean = mean(theta_s, na.rm=T),
    theta_s_median = median(theta_s, na.rm=T),
  ),
  rg_w_theta %>% group_by(MAG, reads_source, lifestyle) %>% filter(coverage > 5) %>% filter(theta_ns != Inf) %>%
  filter((start %% 2 ) == 1) %>%
  summarise(
    theta_ns_mean = mean(theta_ns, na.rm=T),
    theta_ns_median = median(theta_ns, na.rm=T),
    n = n()
  ),
  by = join_by(MAG, reads_source, lifestyle)
  
)

```
```{r}
rg_w_theta_sum %>% ggplot(aes(x=theta_s_mean, y=log(theta_ns_mean))) + geom_point(aes(size=log10(n))) + labs(y="Natural Log \U03B8N / \U03B8S ", x="\U03B8S", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + theme_bw() + geom_smooth(method="lm", se=F, color="black")

```


```{r}
rg_w_theta_sum %>% ggplot(aes(x=log(theta_s_mean), y=log(theta_ns_mean))) + geom_point(aes(size=log10(n))) + labs(y="Natural Log \U03B8N / \U03B8S ", x="Natural Log \U03B8S", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + theme_bw() + geom_smooth(method="lm", se=F, color="black")

rg_w_theta_sum_eo %>% ggplot(aes(x=log(theta_s_mean), y=log(theta_ns_mean))) + geom_point(aes(size=log10(n))) + labs(y="Natural Log \U03B8N / \U03B8S Odd positions ", x="Natural Log \U03B8S Even positions", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + theme_bw() + geom_smooth(method="lm", se=F, color="black")
```

```{r}
rg_w_theta_sum_eo %>% ggplot(aes(x=log(theta_s_mean), y=log(theta_ns_mean), color=factor(reads_source))) + geom_point(aes(size=log10(n))) + labs(y="Natural Log \U03B8N / \U03B8S Odd positions ", x="Natural Log \U03B8S Even positions", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + theme_bw() + geom_smooth(method="lm", se=F)
```


```{r}
rg_w_theta_sum %>% ggplot(aes(x=theta_s_mean, y=log(theta_ns_mean), size=log10(n))) + geom_point() + labs(y="Natural Log ( \U03B8N / \U03B8S ) ", x="\U03B8S", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + facet_wrap(~reads_source) 
```

```{r}
 rg_w_theta_sum %>% group_by(reads_source) %>% summarise(p_val = summary(lm(data=cur_data(), formula=log(theta_ns_mean) ~ theta_s_mean))[["coefficients"]][2,4],r2 = summary(lm(data=cur_data(), formula=log(theta_ns_mean) ~ theta_s_mean))$r.squared)
```

```{r}
rg_w_theta_sum %>% ggplot(aes(x=theta_s_mean, y=log(theta_ns_mean), size=log10(n), shape=factor(reads_source), color=MAG)) + geom_point() + labs(y="Natural Log ( \U03B8N / \U03B8S ) ", x="\U03B8S", size="Log10 \n # of genes") + theme(text=element_text(size=22)) + facet_wrap(~lifestyle) 

```
Gillaume plot
```{r}
rg_w_theta %>% group_by(MAG, reads_source) %>% filter(gene_length > 500 & gene_length < 2000) %>% filter(coverage > 5) %>% summarise(mag_pi_S = sum(SNV_S_count), mag_pi_N = sum(SNV_N_count), n=n()) %>% ggplot(aes(x=mag_pi_S, y=mag_pi_N)) + geom_point(aes(size=log10(n), color=MAG, shape=factor(reads_source))) + labs(x="MAG S count",y="MAG N count",size="log10 # of genes") + theme(text=element_text(size=24)) + theme_bw() + coord_fixed()
```

Just a brief experiment into if we can cluster genes into nice clusters with multiple members (for most, not really)
```{r}
annotations_M5 <- read_delim("~/Projects/Sediments/all_depths/combined_annotations.txt",delim="\t",comment="#")
head(annotations_M5)
```
```{r}
annotations_M5 %>% summarise(n(), sum(Preferred_name != "-"), sum(KEGG_ko != "-"), sum(PFAMs != "-"))
```



How many genes per MAG per depth are under positive selection?  The threshold here is up for some debate.  IS doesn't report pNpS if there's not sufficient variants so I just use whether they report pNpS as a crude filter 


```{r}
left_join(rg_w_theta, annotations_M5, join_by(gene == query)) %>% filter((KEGG_ko != "-") & (!is.na(pNpS_variants)) & (coverage > 5)) %>% ggplot(aes(x=theta_ns)) + geom_histogram(bins=100) + xlim(0,5)
```
```{r}
left_join(rg_w_theta, annotations_M5, join_by(gene == query)) %>% filter((KEGG_ko != "-") & (!is.na(pNpS_variants)) & (coverage > 5)) %>% ggplot(aes(x=theta_ns)) + geom_histogram(binwidth=0.05) + xlim(0.5,2) + geom_vline(xintercept = 1.4, linetype="dotted",color="red")
```

```{r}
genes_under_positive_selection_M5 <- left_join(rg_w_theta, annotations_M5, join_by(gene == query)) %>% filter((KEGG_ko != "-") & (!is.na(pNpS_variants)) & (coverage > 5))

genes_under_positive_selection_M5 <- genes_under_positive_selection_M5 %>% mutate(pos_sel = theta_ns > 1.4) %>% group_by( MAG, reads_source) %>% summarise(n_genes = n(),  n_pos_sel = sum(pos_sel, na.rm=T)) %>% mutate(pos_sel_fr = n_pos_sel / n_genes )
```



```{r}
left_join(genes_under_positive_selection_M5 , raw_genome_M5_3 %>% select(genome, length) %>% distinct(genome, length) %>% mutate(MAG = gsub("refs/","", genome)), join_by(MAG == MAG)) %>% group_by( MAG, reads_source) %>% ggplot(aes(x=factor(reads_source), y=pos_sel_fr, color=log10(length))) + geom_point() + facet_wrap(~MAG) + theme_bw()
```











```{r}
lill_tax <- read.table("~/Downloads/raw_data_16s/Lilian_ASV_table_taxonomy.txt",sep="\t",header=T)
lill_tax <- lill_tax %>% mutate(domain_phyla = paste0(DOMAIN,"_",Phylum))
head(lill_tax)
```
```{r}
lill_tax2 <- lill_tax %>%
  pivot_longer(cols = starts_with("M"),names_to = "name", values_to = "count") %>%
  separate_wider_delim(cols="name",delim="_",names=c("Site","Replicate","Depth","Type")) %>% 
  mutate(Depth = as.numeric(gsub("x",".", Depth))) %>% 
  mutate(domain_phyla = ifelse(DOMAIN == "Archaea", "All Archaea",Phylum))
```

```{r}
top_phyla <- lill_tax2 %>% filter(Site == "M5",Type =="DNA") %>% group_by(domain_phyla) %>%  summarise(dna_sum = sum(count)) %>% top_n( 11)
read_totals <- lill_tax2 %>% group_by(Depth, Type, Site, Replicate) %>% summarise(read_total = sum(count))
lill_tax2 <- left_join(lill_tax2, read_totals)

color_brewer_w_gray <- c("#a6cee3","#1f78b4","#b2df8a","#33a02c","#fb9a99","#e31a1c","#fdbf6f","#aaaaaa","#ff7f00","#cab2d6","#6a3d9a","#ffff99","#b15928")
color_brewer <- c("#a6cee3","#1f78b4","#b2df8a","#33a02c","#fb9a99","#e31a1c","#fdbf6f","#ff7f00","#cab2d6","#6a3d9a","#ffff99","#b15928")
lill_tax2 %>% filter((Site == "M5") & (Type == "RNA") & (domain_phyla %in% top_phyla$domain_phyla)) %>% group_by(domain_phyla,Depth) %>% summarise(rna_count = sum(count)) %>% uncount(rna_count) %>% ggplot(aes(x=Depth,fill=domain_phyla)) + geom_bar() + labs(x="Cm Below Seafloor",y="16S RNA read count",fill="Phylum") + theme_bw() + scale_fill_manual(values=color_brewer) + geom_vline(xintercept = 6.5, linetype="dotted",color="black") + theme(text=element_text(family="Helvetica",size=22)) 

```

```{r}
lill_tax2 %>% mutate(domain_phyla = ifelse(domain_phyla %in% top_phyla$domain_phyla, domain_phyla, "Other Bacteria")) %>% filter((Site == "M5") & (Type == "RNA") & ((domain_phyla %in% top_phyla$domain_phyla)|(domain_phyla == "Other Bacteria"))) %>% group_by(domain_phyla,Depth, read_total) %>% summarise(phyla_reads = sum(count)) %>% mutate(phyla_relabs = phyla_reads / read_total) %>% ggplot(aes(x=Depth,y=phyla_relabs, fill=domain_phyla)) + geom_bar(stat="identity") + labs(x="Cm Below Seafloor",y="16S RNA rel. abd.",fill="Phylum") + theme_bw() + scale_fill_manual(values=color_brewer_w_gray) + geom_vline(xintercept = 6.5, linetype="dotted",color="black") + theme(text=element_text(family="Helvetica",size=22)) 
```

