---
title: "new_figures"
output: html_document
date: "2025-06-19"
---

```{r}
library(tidyverse)
library(gt) #seems to be the best table library unless you need it as an image in which case ggpubr::ggtexttable
```



```{r}
mag_table <- read_delim("~/Projects/Sediments/MAG_table_2.tsv") 
mag_table %>%
  mutate(MAG = str_extract(MAG, "M[0-9]_[0-9]+.[0-9]+")) %>%
  select(-Species) %>%
  gt() %>%
  tab_footnote(
    footnote = "This is the only MAG annotated with a species: JAAXQD01 sp012974475",
    locations = cells_body(columns = "Genus", rows = 15)
    ) %>% opt_stylize( style = 2, color = "gray", add_row_striping = TRUE)
```



these two are generated kable_classic_2()these two are generated in making_gene_dataframes.Rmd
```{r}
popgen_plus_anno_m5 <- read_delim("~/Projects/Sediments/MAG_popgen_plus_anno_M5.tsv") %>% mutate(scaffold = str_extract(gene, "contig_[0-9]+")) %>% filter(MAG != "M5_3.64.filtered.fa", scaffold != "contig_4729")
popgen_plus_anno_m1 <- read_delim("~/Projects/Sediments/MAG_popgen_plus_anno_M1.tsv") %>% mutate(scaffold = str_extract(gene, "contig_[0-9]+"))
```

```{r}
pnps_all <- popgen_plus_anno_m5 %>%
  filter(!is.infinite(p_NS)) %>%
  ggplot(aes(x=log10(coverage + 0.01), y=log10(p_NS+ 0.01))) +
    geom_point() +
    geom_smooth(method="lm") +
    theme_bw() +
    theme(text=element_text(size=22, family="Helvetica")) +
    labs(y="log10(pN/pS + 0.01)")
pnps_5plus <- popgen_plus_anno_m5 %>%
  filter(!is.infinite(p_NS)) %>%
  filter(coverage >= 5)  %>%
  ggplot(aes(x=log10(coverage + 0.01), y=log10(p_NS+ 0.01))) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_bw() +
  theme(text=element_text(size=22, family="Helvetica")) +
  labs(y="log10(pN/pS + 0.01)")
theta_all <- popgen_plus_anno_m5 %>%
  filter(!is.infinite(theta_NS)) %>%
  ggplot(aes(x=log10(coverage+0.01), y=log10(theta_NS+ 0.01))) +
    geom_point() +
    geom_smooth(method="lm" ) +
    theme_bw() +
    theme(text=element_text(size=22, family="Helvetica")) +
    labs(y="log10(\u03B8N/\u03B8S + 0.01)")
theta_5plus <- popgen_plus_anno_m5 %>%
  filter(!is.infinite(theta_NS)) %>%
  filter(coverage >= 5)  %>%
  ggplot(aes(x=log10(coverage+0.01), y=log10(theta_NS+ 0.01))) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_bw() +
  theme(text=element_text(size=22, family="Helvetica")) +
  labs(y="log10(\u03B8N/\u03B8S + 0.01)")

supl_fig_1 <- cowplot::plot_grid(pnps_all, pnps_5plus, theta_all, theta_5plus, labels=c("pN/pS ~ cvg", "pN/pS ~ cvg ≥ 5x", "tN/tS ~ cvg", "tN/tS ~ cvg ≥ 5x"), ncol=2)
supl_fig_1
```
see reference on behavior of theta,  +

```{r}
summary(popgen_plus_anno_m5 %>% filter(!is.infinite(p_NS)) %>% lm(data = ., formula = log10(p_NS + 0.01) ~ log10(coverage+0.01)))
```
```{r}
summary(popgen_plus_anno_m5 %>% filter(!is.infinite(p_NS)) %>% filter(coverage >= 5) %>% lm(data = ., formula = log10(p_NS + 0.01) ~ log10(coverage + 0.01)))
```
```{r}
summary(popgen_plus_anno_m5 %>% filter(!is.infinite(theta_NS)) %>% lm(data = ., formula = log10(theta_NS + 0.01) ~ log10(coverage+0.01)))

```
```{r}
summary(popgen_plus_anno_m5 %>% filter(!is.infinite(theta_NS)) %>% filter(coverage >= 5) %>% lm(data = ., formula = log10(theta_NS + 0.01) ~ log10(coverage+0.01)))

```


```{r}
generate_hex_plot_pi <- function(plot_site, plot_lifestyle, plot_palette, plot_title, plot_df){
  ph = sequential_hcl(2, palette=plot_palette)[1] # grab the high and low colors from the colorspace library
  pl = sequential_hcl(2, palette=plot_palette)[2]
  return(#there is 0 need to save this object inside the function
    plot_df %>%
      filter((lifestyle == plot_lifestyle) & (site == plot_site)) %>%
      ggplot(aes(x=pi_NS_25, y=pi_NS_3)) +
        geom_hex() +
        labs(x="πN/πS 25 CmBSF", y="πN/πS 3 CmBSF",title=plot_title) +
        xlim(0,1) +
        ylim(0,1) +
        coord_fixed() + 
        geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
        scale_fill_continuous(high=ph,low=pl) +# setting the color with aes fill would set each hex the same and remove density info
        theme_bw() + 
        theme(text=element_text(family="Helvetica", size=22))
  )
}
```

```{r}
combined_genes_3_25 <- rbind(popgen_plus_anno_m5 %>% mutate(site="M5"), popgen_plus_anno_m1 %>% mutate(site="M1")) %>% filter((reads_source == 25) |(reads_source == 3)) %>% filter(coverage >= 5) %>% group_by(gene, reads_source) %>% slice_head(n=1) %>% ungroup() %>% select(reads_source, pi_NS, theta_NS, genome, lifestyle, site, gene ) %>% pivot_wider( names_from = reads_source, values_from = c(pi_NS, theta_NS)) %>% filter(!is.na(pi_NS_3) & !is.na(pi_NS_25))
```


```{r}
m5_sp_p <- generate_hex_plot_pi("M5", "Surface Specialist","RedOr","M5 Surface Spec.",combined_genes_3_25)
m5_gen_p <- generate_hex_plot_pi("M5", "Generalist","PurpOr","Generalist",combined_genes_3_25)
m5_ssp_p <- generate_hex_plot_pi("M5", "Subsurface Specialist","Purp","Subsurf. Spec.",combined_genes_3_25)
m1_sp_p <- generate_hex_plot_pi("M1", "Surface Specialist","RedOr","M1",combined_genes_3_25)
m1_gen_p <- generate_hex_plot_pi("M1", "Generalist","PurpOr","",combined_genes_3_25)
m1_ssp_p <- generate_hex_plot_pi("M1", "Subsurface Specialist","Purp","",combined_genes_3_25)
```

```{r}
m5_row_p <- cowplot::plot_grid(m5_sp, m5_gen, m5_ssp, nrow=1)
m1_row_p <- cowplot::plot_grid(m1_sp, m1_gen, m1_ssp, nrow=1)
panel_2_pi <- cowplot::plot_grid(m5_row,m1_row,ncol=1)
```
```{r}
combined_genes_3_25 %>% group_by(site, lifestyle) %>% summarise(n = n())
```


```{r}
generate_hex_plot_theta <- function(plot_site, plot_lifestyle, plot_palette, plot_title, plot_df){
  ph = sequential_hcl(2, palette=plot_palette)[1] # grab the high and low colors from the colorspace library
  pl = sequential_hcl(2, palette=plot_palette)[2]
  return(#there is 0 need to save this object inside the function
    plot_df %>%
      filter((lifestyle == plot_lifestyle) & (site == plot_site)) %>%
      ggplot(aes(x=theta_NS_25, y=theta_NS_3)) +
        geom_hex() +
        labs(x="\u03B8N/\u03B8S 25 CmBSF", y="\u03B8N/\u03B8S 3 CmBSF",title=plot_title) +
        xlim(0,1) +
        ylim(0,1) +
        coord_fixed() + 
        geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
        scale_fill_continuous(high=ph,low=pl) +# setting the color with aes fill would set each hex the same and remove density info
        theme_bw() + 
        theme(text=element_text(family="Helvetica", size=22))
  )
}
```

```{r}
m5_sp_t <- generate_hex_plot_theta("M5", "Surface Specialist","RedOr","M5 Surface Spec.",combined_genes_3_25)
m5_gen_t <- generate_hex_plot_theta("M5", "Generalist","PurpOr","Generalist",combined_genes_3_25)
m5_ssp_t <- generate_hex_plot_theta("M5", "Subsurface Specialist","Purp","Subsurf. Spec.",combined_genes_3_25)
m1_sp_t <- generate_hex_plot_theta("M1", "Surface Specialist","RedOr","M1",combined_genes_3_25)
m1_gen_t <- generate_hex_plot_theta("M1", "Generalist","PurpOr","",combined_genes_3_25)
m1_ssp_t <- generate_hex_plot_theta("M1", "Subsurface Specialist","Purp","",combined_genes_3_25)
```

```{r}
m5_row_t <- cowplot::plot_grid(m5_sp_t, m5_gen_t, m5_ssp_t, nrow=1)
m1_row_t <- cowplot::plot_grid(m1_sp_t, m1_gen_t, m1_ssp_t, nrow=1)
panel_2_theta <- cowplot::plot_grid(m5_row_t,m1_row_t,ncol=1)
```

```{r}
rbind(popgen_plus_anno_m5 %>% mutate(site = "M5"), popgen_plus_anno_m1 %>% mutate(site = "M1")) %>%
  ggplot(aes(x=factor(reads_source), y = log10(theta_NS+ 0.001), color=is_busco)) + 
  geom_violin(draw_quantiles = c(0.5)) + 
  facet_wrap(~site) + 
  labs(x = "CmBSF", y = "Log10 \u03B8N/\u03B8S + 0.001", color="Gene is BUSCO?") + 
  theme_bw() + 
  theme(text=element_text(family="Helvetica", size=22))
```



```{r}
all_depths_pi_n_pi_s <- rbind(popgen_plus_anno_m5 %>% mutate(site = "M5"), popgen_plus_anno_m1 %>% mutate(site = "M1") )
all_depths_pi_s <- all_depths_pi_n_pi_s %>%
  group_by(MAG, reads_source, lifestyle) %>%
      mutate(number = as.integer(str_extract(gene, "(?<=_)[0-9]+$"))) %>%
      filter(coverage > 5) %>%
      filter((number %% 2 ) == 0) %>% #even numbered starts -> theta-hat S, odd numbered starts -> theta-hat NS
  mutate(
    cn_p_N = n_variants_N / (Ln * coverage), # coverage normalized pN
    cn_p_S = n_variants_S / (Ls * coverage)
    ) %>%
  mutate(
    cn_p_NS = cn_p_N / cn_p_S
  ) %>% 
  filter(
    (n_variants_S > 0) & (!is.na(n_variants_N)) & (pi_S > 0) & (theta_S > 0) & !is.infinite(theta_S) & (coverage > 5)
    ) %>%
  group_by(MAG, reads_source, site) %>% 
  summarise(
#    ln_pi_ns = mean(log(pi_NS + 0.0001)),
    ln_pi_s = mean(log(pi_S)),
 #   ln_theta_ns = mean(log(theta_NS+ 0.0001)),
    ln_theta_s = mean(log(theta_S)),
  #  ln_p_ns = mean(log(p_NS+ 0.0001)),
    ln_p_s = mean(log(p_S)),
  #  ln_cn_p_ns = mean(log(cn_p_NS+ 0.0001)),
    ln_cn_p_s = mean(log(cn_p_S)),
    coverage = mean(coverage),
    n_s = n()
  )
all_depths_pi_ns <- all_depths_pi_n_pi_s %>%
  group_by(MAG, reads_source, lifestyle) %>%
      mutate(number = as.integer(str_extract(gene, "(?<=_)[0-9]+$"))) %>%
      filter(coverage > 5) %>%
      filter((number %% 2 ) != 0) %>% #even numbered starts -> theta-hat S, odd numbered starts -> theta-hat NS
  mutate(
    cn_p_N = n_variants_N / (Ln * coverage), # coverage normalized pN
    cn_p_S = n_variants_S / (Ls * coverage)
    ) %>%
  mutate(
    cn_p_NS = cn_p_N / cn_p_S
  ) %>% 
  filter(
    (n_variants_S > 0) & (!is.na(n_variants_N)) & (pi_S > 0) & (theta_S > 0) & !is.infinite(theta_S) & (coverage > 5)
    ) %>%
  group_by(MAG, reads_source, site) %>% 
  summarise(
    ln_pi_ns = mean(log(pi_NS + 0.0001)),
#    ln_pi_s = mean(log(pi_S)),
    ln_theta_ns = mean(log(theta_NS+ 0.0001)),
 #   ln_theta_s = mean(log(theta_S)),
    ln_p_ns = mean(log(p_NS+ 0.0001)),
  #  ln_p_s = mean(log(p_S)),
    ln_cn_p_ns = mean(log(cn_p_NS+ 0.0001)),
  #  ln_cn_p_s = mean(log(cn_p_S)),
    n_ns = n()
  )
all_depths_pi_n_pi_s_summary <- left_join(all_depths_pi_s, all_depths_pi_ns) %>% mutate(n = n_s + n_ns)
```

```{r}
panel_3_pinpis = all_depths_pi_n_pi_s_summary %>%
  ggplot(aes(x=ln_pi_s, y = ln_pi_ns, color=factor(reads_source),  size=log10(n))) +
    geom_point() +
    geom_smooth(method="lm", se=F) +
    labs(x="Nat. log πS", y="Nat. log πN/πS") + 
    facet_wrap(~site) + 
    theme_bw() +
    theme(text=element_text(family="Helvetica", size=22))
panel_3_pinpis
```
```{r}
all_depths_pi_n_pi_s_summary %>% 
  group_by(site, reads_source) %>%
  summarise(
    r2 = summary(lm(data=cur_data(), ln_pi_ns ~ ln_pi_s))$r.squared,
    p_val = summary(lm(data=cur_data(), ln_pi_ns ~ ln_pi_s))[["coefficients"]][2,4]
    )
```


```{r}
panel_3_thetan_thetas = all_depths_pi_n_pi_s_summary %>%
  ggplot(aes(x=ln_theta_s, y = ln_theta_ns, size=log10(n))) +
    geom_point(aes(color=factor(reads_source))) +
    geom_smooth(aes(color=factor(reads_source)),method="lm", se=F) +
    geom_smooth(method="lm", se=F, linetype="dashed", color="black") + 
    labs(x="Nat. log \u03B8S", y="Nat. log \u03B8N/\u03B8S", color="CmBSF", size="Log10 # of genes") + 
    facet_wrap(~site) + 
    theme_bw() +
    theme(text=element_text(family="Helvetica", size=22))
panel_3_thetan_thetas
```
```{r}
regression_tnts <- all_depths_pi_n_pi_s_summary %>% 
  rename(CmBSF = reads_source) %>% 
  group_by(site, CmBSF) %>%
  summarise(
    slope = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))[["coefficients"]][2,1],
    r2 = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))$r.squared,
    p_val = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))[["coefficients"]][2,4]
    ) %>% ungroup()
regression_tnts <- rbind(regression_tnts,
                         all_depths_pi_n_pi_s_summary %>% 
                           mutate(reads_source = "All") %>%
                           rename(CmBSF = reads_source) %>%
                            group_by(site, CmBSF) %>%
                            summarise(
                              slope = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))[["coefficients"]][2,1],
                              r2 = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))$r.squared,
                              p_val = summary(lm(data=cur_data(), ln_theta_ns ~ ln_theta_s))[["coefficients"]][2,4]
                              ) %>% ungroup()
                         )
regression_table_tnts <- ggpubr::ggtexttable(regression_tnts, theme=ggpubr::ttheme("lBlackWhite"), rows=NULL)
regression_table_tnts
```
does coverage contribute?
```{r}
summary(all_depths_pi_n_pi_s_summary %>% filter(site == "M5") %>% lm(data=., formula = ln_theta_ns ~ ln_theta_s + log(coverage)))
```
```{r}
summary(all_depths_pi_n_pi_s_summary %>% filter(site == "M5") %>% lm(data=., formula = ln_theta_ns ~ ln_theta_s ))
```



```{r}
bmcs <-  popgen_plus_anno_m5 %>% filter(is_busco) %>% group_by(MAG, reads_source) %>% summarise(bmc = mean(coverage, na.rm=T)) %>% ungroup()
gene_corr_coefs <- left_join(
  popgen_plus_anno_m5 %>%
    select(MAG, gene, reads_source, coverage) %>%
    pivot_wider(
      id_cols=c("MAG","gene"),
      names_from=reads_source,
      values_from=c(coverage),
      names_prefix = "coverage_") %>%
    pivot_longer(
      cols=starts_with("coverage_"),
      names_to="reads_source",
      values_to = "coverage") %>%
    mutate(
      coverage = ifelse(is.na(coverage),0,coverage),
      reads_source=as.integer(gsub("coverage_","",reads_source))),
  bmcs) %>%
  group_by(MAG, gene) %>%
  #slice_head(n=10)
  summarise(corr = cor(coverage, bmc), covg_var = var(coverage / bmc))

gene_corr_coefs <- gene_corr_coefs %>% mutate(is_busco = gene %in% busco_gene_ids_M5$gene)
head(gene_corr_coefs)
```


```{r}
core_acc_evol_table <- left_join(popgen_plus_anno_m5, gene_corr_coefs) %>%
  filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
  mutate(pang_status = ifelse(corr >= 0.75, "Core", "Acc.")) %>% #pangenome status
  mutate(evol_status = ifelse(theta_NS < 0.838, "Conserved", ifelse(theta_NS < 1.4, "Neutral", "Positive"))) %>%
  group_by(genome, pang_status, evol_status, reads_source) %>%
  summarise(n = n()) %>% 
  ungroup() 
core_acc_evol_table %>%
  pivot_wider(values_from=n, names_from = c(pang_status, evol_status))
```
```{r}
ggpubr::ggtexttable(core_acc_evol_table %>%
                      rename(CmBSF = reads_source) %>%
                      pivot_wider(values_from=n, names_from = c(pang_status, evol_status), names_sep = " ") ,
 theme=ggpubr::ttheme("lBlackWhite"), rows=NULL)

```




here I work on panel 4, first section is function declarations





Determine what genes are environmentally sensitive (this is here because it's imperative it uses the same popgen_plus_anno df that the other analyses do)

```{r}
clusters_M5 <- read_delim("~/Projects/Sediments/clustered_aaai50_cluster_M5.tsv",col_names = c("cluster_rep","cluster_member"),delim="\t")
clusters_M5 <- clusters_M5 %>% group_by(cluster_rep) %>% mutate(cluster_number = cur_group_id())

ppa_m5_clusters <- left_join(popgen_plus_anno_m5, clusters_M5 %>% select(cluster_member, cluster_number), join_by(gene == cluster_member))

ppa_m5_clusters <- ppa_m5_clusters %>%
  select(MAG, gene, cluster_number, is_busco, coverage, reads_source) %>%
  pivot_wider(values_from=coverage, names_from = reads_source,names_prefix="coverage_") %>% 
  pivot_longer(cols=starts_with("coverage"),names_to = "reads_source",values_to="coverage",names_prefix="coverage_",values_drop_na = FALSE) %>%
  mutate(coverage = ifelse(is.na(coverage), 0, coverage)) 
busco_MAG_avg_M5 <- ppa_m5_clusters %>% filter(is_busco) %>%  group_by(MAG, reads_source) %>% summarise(n = n(), bmc = mean(coverage, na.rm=T))


ppa_m5_clusters <- left_join(ppa_m5_clusters, busco_MAG_avg_M5, join_by(MAG, reads_source)) %>%
  filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
  mutate(reads_source = as.integer(reads_source)) %>%
  mutate(log10_fold_change = log10((coverage + 0.001) / bmc)) %>%
  mutate(envt = factor(ifelse(reads_source < 7, "BT","SRZ"))) #here I define an environment, other versions of this analysis have done a regression based on log10 CmBSF

reg_df_m5 <- ppa_m5_clusters %>%
  group_by(cluster_number) %>%
  reframe(
    linear_model_summary = summary(lm(log10_fold_change ~ envt))[["coefficients"]][2,]
  )

reg_df_m5$value_name = rep(c("coef","SE","t_val","p_val"),length(reg_df_m5$cluster_number)/4)

reg_df_m5 <- reg_df_m5 %>%
  pivot_wider(values_from = linear_model_summary, names_from = value_name) 

ppa_m5_clusters <- left_join(ppa_m5_clusters, reg_df_m5, join_by(cluster_number)) %>% filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa"))
```
```{r}
m5_gene_heatmap <- ppa_m5_clusters %>%
  arrange(desc(coef)) %>%
  ggplot(aes(x=factor(reads_source), y=gene, fill=log10_fold_change)) +
  geom_tile() +
  facet_wrap(~MAG, ncol=2, scales="free_y") + 
  colorspace::scale_fill_binned_diverging(palette = "Tropic") +
  theme_bw() + 
  theme(axis.text.y = element_blank(), text=element_text(family="Helvetica", size=22))+
  labs(x = "CmBSF", fill="Relative frequency") 
m5_gene_heatmap
```





everything below this has been dropped :(, too bandaid-y and not really solid enough for primetime.  Could come back to it with more persisters??  Need to find a way around the read stealing problem though.  Ideal bacteria for this analysis has lots of within MAG diversity but is very diverged from the rest of the community, also needs to be a solid MAG >90% complete with high N50






```{r}
#one of the contigs in M5_3.252 is on the whole very depth sensitive so I remove it from this plot 
m5_pvals_plot <- ppa_m5_clusters %>%
  distinct(cluster_number, .keep_all=T) %>%
  mutate(genome = gsub(".filtered.fa","",MAG)) %>%
  mutate(contig = gsub("_[0-9]+?$","", gene)) %>%
  #filter(contig != "contig_4729") %>%
  ggplot(aes(x=p_val)) +
    geom_histogram(binwidth = 0.03) +
    facet_wrap(~genome) +
    theme_bw() +
    labs(x="P-value")

m5_pvals_plot
```

```{r}
lowest_pvals_m5 <- ppa_m5_clusters %>% filter(MAG == "M5_3.252.filtered.fa") %>% group_by(cluster_number) %>% slice_head(n=1) %>% ungroup() %>% slice_min(p_val , n=5)
lowest_pvals_m5
```


```{r}
gene_by_gene_252 <- ppa_m5_clusters %>%
  filter(gene %in% lowest_pvals_m5$gene) %>%
  ggplot(aes(y=log10_fold_change, x=factor(reads_source))) +
    geom_point() +
    geom_smooth(method="lm",se=F) +
    facet_wrap(~gene, scales = "free_y") +
    theme_bw() +
    labs(x="CmBSF", y="Log10 BUSCO normed. covg.")  
gene_by_gene_252
```
```{r}
gene_by_gene_fxs <- ggpubr::ggtexttable(
  popgen_plus_anno_m5 %>%
    group_by(gene) %>%
    slice_head(n=1) %>%
    ungroup() %>%
    filter(gene %in% lowest_pvals_m5$gene) %>%
    mutate(Entry = ifelse(is.na(Entry), Phrog, Entry)) %>% 
    select(gene, Entry, `Protein names`, `Gene Names`),
  theme=ggpubr::ttheme("lBlackWhite"), rows=NULL)
gene_by_gene_fxs
```

```{r}
top_5_pos <- ppa_m5_clusters %>%
  filter((coef > 0) & p_val < 0.1, MAG == "M5_3.252.filtered.fa") %>%
  group_by(gene) %>%
  slice_head(n= 1) %>%
  ungroup() %>%
  slice_min(tibble(p_val, coef), n=5, with_ties = T)

ppa_m5_clusters %>%
  filter(gene %in% top_5_pos$gene)%>%
  ggplot(aes(y=log10_fold_change, x=factor(reads_source))) +
    geom_point() +
    geom_smooth(method="lm",se=F) +
    facet_wrap(~gene, scales = "free_y") +
    theme_bw() +
    labs(x="log10 CmBSF", y="Log10 BUSCO normed. covg.")
```
```{r}
ggpubr::ggtexttable(
  popgen_plus_anno_m5 %>%
    group_by(gene) %>%
    slice_head(n=1) %>%
    ungroup() %>%
    filter(gene %in% top_5_pos$gene) %>%
    mutate(Entry = ifelse(is.na(Entry), Phrog, Entry)) %>% 
    select(gene, Entry, `Protein names`, `Gene Names`),
  theme=ggpubr::ttheme("lBlackWhite"), rows=NULL)
```





```{r}
clusters_M1 <- read_delim("~/Projects/Sediments/clustered_aaai50_cluster_M1.tsv",col_names = c("cluster_rep","cluster_member"),delim="\t")
clusters_M1 <- clusters_M1 %>% group_by(cluster_rep) %>% mutate(cluster_number = cur_group_id())

ppa_m1_clusters <- left_join(popgen_plus_anno_m1, clusters_M1 %>% select(cluster_member, cluster_number), join_by(gene == cluster_member))

ppa_m1_clusters <- ppa_m1_clusters %>%
  select(MAG, gene, cluster_number, is_busco, coverage, reads_source) %>%
  pivot_wider(values_from=coverage, names_from = reads_source,names_prefix="coverage_") %>% 
  pivot_longer(cols=starts_with("coverage"),names_to = "reads_source",values_to="coverage",names_prefix="coverage_",values_drop_na = FALSE) %>%
  mutate(coverage = ifelse(is.na(coverage), 0, coverage)) 
busco_MAG_avg_M1 <- ppa_m1_clusters %>% filter(is_busco) %>%  group_by(MAG, reads_source) %>% summarise(n = n(), bmc = mean(coverage, na.rm=T))


ppa_m1_clusters <- left_join(ppa_m1_clusters, busco_MAG_avg_M1, join_by(MAG, reads_source)) %>%
  filter(MAG %in% c("refs/M1_3.182.filtered.fa","refs/M1_3.148.filtered.fa")) %>%
  mutate(reads_source = as.integer(reads_source)) %>%
  mutate(log10_fold_change = log10((coverage + 0.001) / bmc)) %>% 
  mutate(envt = ifelse(reads_source < 25, "BT", "SRZ"))

reg_df_m1 <- ppa_m1_clusters %>%
  group_by(cluster_number) %>%
  reframe(
    linear_model_summary = summary(lm(log10_fold_change ~log10(reads_source)))[["coefficients"]][2,]
  )

reg_df_m1$value_name = rep(c("coef","SE","t_val","p_val"),length(reg_df_m1$cluster_number)/4)

reg_df_m1 <- reg_df_m1 %>%
  pivot_wider(values_from = linear_model_summary, names_from = value_name) 

ppa_m1_clusters <- left_join(ppa_m1_clusters, reg_df_m1, join_by(cluster_number)) 
```
```{r}
combined_gene_heatmap <- rbind(ppa_m5_clusters, ppa_m1_clusters) %>%
  mutate(genome = str_extract(MAG, "M[15]_[0-9]+.[0-9]+")) %>%
  select(gene, genome, reads_source, log10_fold_change) %>%
  pivot_wider(names_from = reads_source, values_from = log10_fold_change, names_prefix = "l10_fc_") %>%
  arrange(desc(l10_fc_25)) %>%
  pivot_longer(matches(".*_[0-9]+"), names_to="reads_source", values_to="log10_fold_change", names_prefix = "l10_fc_") %>%
  mutate(reads_source = as.integer(reads_source)) %>%
  ggplot(aes(x=factor(reads_source), y=gene, fill=log10_fold_change)) +
    geom_tile() +
    facet_wrap(~genome, ncol=2, scales="free_y") + 
    colorspace::scale_fill_binned_diverging(palette = "Tropic") +
    theme_bw() + 
    theme(axis.text.y = element_blank(), text=element_text(family="Helvetica", size=22))+
    labs(x = "CmBSF", fill="Log10 Relative frequency") 
combined_gene_heatmap
```



```{r}
ggpubr::ggtexttable(
rbind(ppa_m5_clusters, ppa_m1_clusters) %>% 
  group_by(MAG) %>%
  summarise(
    `True Pos Frac.` = 1 - qvalue::qvalue(p = p_val, fdr.level = 0.1)$pi0
  ),
theme=ggpubr::ttheme("lBlackWhite"), rows=NULL)
```


```{r}
#one of the contigs in M5_3.252 is on the whole very depth sensitive so I remove it from this plot 
m1_pvals_plot <- ppa_m1_clusters %>%
  distinct(cluster_number, .keep_all=T) %>%
  mutate(genome = gsub(".filtered.fa","",MAG)) %>%
  mutate(contig = gsub("_[0-9]+?$","", gene)) %>%
  #filter(contig != "contig_4729") %>%
  ggplot(aes(x=p_val)) +
    geom_histogram(binwidth = 0.03) +
    facet_wrap(~genome) +
    theme_bw() +
    labs(x="P-value")

m1_pvals_plot
```

```{r}
sim_data = rbind(data.frame(p_val = runif(4000), source="H0"), data.frame(p_val = rbeta(400, 1, 10), source="HA"))
sim_pvals_plot <- sim_data %>% ggplot(aes(x=p_val, fill=source)) + geom_histogram() + ggthemes::scale_fill_colorblind() + labs(x="P-value") + 
theme(legend.position="none", text=element_text(size=14)) 

cowplot::plot_grid(cowplot::plot_grid(sim_pvals_plot, m1_pvals_plot,rel_widths = c(1,2), rows = 1), m5_pvals_plot, cols=1)
```



```{r}
scaffold_info_m5 = read_delim("~/Projects/Sediments/IS_raw_M5_3_3_scaffold_info.tsv") %>% select(scaffold, length) %>% arrange(desc(length))
scaffold_info_m1 = read_delim("~/Projects/Sediments/IS_raw_M1_3_3_scaffold_info.tsv") %>% select(scaffold, length) %>% arrange(desc(length))
```

```{r}
make_running_length <- function(genes_df, scaffold_df, mag_name, regression_df){
  rl <- left_join(
    genes_df %>% filter(reads_source==3, MAG==mag_name),
    scaffold_df %>% rename(scaff_length = length)
  ) %>% select(scaffold, scaff_length, MAG) %>%
  group_by(scaffold) %>% slice(1) %>% ungroup() %>%
  arrange(desc(scaff_length)) %>% 
  mutate(scaffold_nb = row_number()) %>%
  mutate(scaff_rl = cumsum(scaff_length))
  rl <- rl %>% mutate(scaff_rl2 = lag(scaff_rl)) %>% mutate(scaff_rl2 = ifelse(is.na(scaff_rl2),0, scaff_rl2)) %>% select(-scaff_rl)
  return(rl)
}
make_zoom_plot <- function(genes_df, scaff_df, myMAG,regression_df, zoom_contig){
  rl = make_running_length(genes_df, scaff_df, myMAG) # running length
  mag_genes_df <- left_join(
                    genes_df %>% filter(MAG == myMAG), 
                    rl ) %>%
                  mutate(running_length = scaff_rl2 + start) %>%
                  rename(ctg_start = scaff_rl2)
  if(!("bmc" %in% colnames(mag_genes_df))){
    busco_MAG_avg <- genes_df %>% filter(is_busco) %>%  group_by(MAG, reads_source) %>% summarise(n = n(), bmc = mean(coverage, na.rm=T))
    mag_genes_df <- left_join(mag_genes_df, busco_MAG_avg)
  }
  
  mag_genes_df <- mag_genes_df %>%
    filter(MAG == myMAG) %>%
    filter((reads_source == 3) | (reads_source == 25)) %>%
    select(gene, MAG, reads_source, coverage, bmc, running_length, ctg_start, scaff_length, scaffold, scaffold_nb, is_busco) %>%
    pivot_wider(names_from = reads_source, values_from = c(coverage,bmc)) %>%
    mutate(delta_c = log10(coverage_3 / bmc_3) - log10(coverage_25 / bmc_25)) %>%
    arrange(desc(scaff_length))
  
  mag_genes_df <- left_join(mag_genes_df, regression_df %>% select(gene, p_val))
  
  #me_terms = 'recombinase|excise|exciase| excisionase|recombination|transposase|lysogenic|prophage|temperate|integrase|repressor'
  #mag_genes_df <- left_join(mag_genes_df, annotation_df) %>% mutate(is_me_gene = grepl(me_terms ,`Protein names`,ignore.case = T))
  mag_genes_df <- mag_genes_df %>% 
   # mutate(status = ifelse(is_busco, "BUSCO", ifelse(p_val < 0.1, ifelse(is_me_gene,"Env. Sens. ME","Env. Sensative"), ifelse(is_me_gene, "ME gene", "Others"))))
    mutate(status = ifelse(is_busco, "BUSCO", ifelse(p_val < 0.1, "Env. Sens.", "Others")))
  
  myplot <- mag_genes_df %>% filter(!is.na(scaffold_nb)) %>% ggplot(aes(x=running_length, y = -delta_c)) +
    geom_point(aes(color=status)) + 
    scale_color_manual(name="Gene Status",values=c("#E16A86", "#00AD9A", "grey")) +
    #scale_color_manual(name="Gene Status",values=c("#E16A86","#909800", "#00AD9A", "#9183E6","grey")) +#colors from colorspace dark3
    ggnewscale::new_scale_color() +
    geom_rug(
      aes(
        x = running_length,
        length=scaff_length,
        color=ifelse(scaffold_nb %% 2 == 0, "even","odd")
        #color=ifelse(scaffold ==zoom_contig, zoom_contig,ifelse(scaffold_nb %% 2 == 0, "even","odd"))
        ), 
      sides="b",
      linewidth=1.5
    ) +
    scale_color_manual(name="scaffolds",values=c("black","grey")) +
    #scale_color_manual(name="scaffolds",values=c("blue","black","grey")) +
    geom_vline(aes(xintercept=ctg_start),linetype="dotted") + 
    ggforce::facet_zoom(xy = scaffold == zoom_contig, horizontal = F) +
    labs(x="Artificial coordinates (bp)",y="Abs. change in Rel. Abd. (3 vs 25 CmBSF)") 
  return(myplot)
}
```

```{r}
zoom_plot_m5_252_3 <-make_zoom_plot(popgen_plus_anno_m5, scaffold_info_m5, "M5_3.252.filtered.fa",ppa_m5_clusters, "contig_1291")
zoom_plot_m5_252_3 
```
```{r}
aligned_br <- cowplot::align_plots(gene_by_gene_252,gene_by_gene_fxs , align='v', axis='t')
bottom_row = cowplot::plot_grid(aligned_br[[1]], aligned_br[[2]],nrow=1, labels = c("B", "C"))
bottom_row
```

```{r}
combined_plot <- cowplot::plot_grid(zoom_plot_m5_252_3, bottom_row, nrow=2, labels= c("A",""))
combined_plot
```




