---
title: "panel_1"
output: html_document
date: "2025-03-12"
---

```{r}
library(ampvis2)
library(tidyverse)
library(extrafont)
font_import()
loadfonts()

theme_set(theme_bw())
theme_update(text=element_text(size=22, family="Helvetica"))

```


```{r}
#Kasper Urup Kjeldsen kode (stolen shamelessly by me)

myotutable <- read.delim("~/Downloads/1ASV_raw_abund_tax.txt")
#1ASV_raw_abund_tax.txt
#1ASV_table_absl_abund_tax.txt

mymetadata <- read.delim("~/Downloads/Lilian_metadata.txt")

d <- amp_load(otutable = myotutable,
              metadata = mymetadata)

nmds_out <- amp_ordinate(d, 
             type = "nmds",
             transform = "none",
             distmeasure = "bray",
             sample_color_by = "Ordering",
             sample_label_by = "sediment_depth_cm",
             sample_label_size = 3,
             filter_species = 0.01,
             sample_colorframe = FALSE,
             sample_shape_by = "Ordering",
             max.overlaps = Inf,
             detailed_output = TRUE) 

nmds_plot <- nmds_out$plot + theme(legend.position="bottom", text=element_text(family="Helvetica", size=16))
nmds_plot
```

```{r}
read_sums <- myotutable %>% rowwise() %>% mutate(read_total = across(ends_with("RNA"), ~sum(.x, na.rm=T))) %>% ungroup() %>% group_by( Genus) %>% summarise(read_total = sum(read_total))
mags_further_analysis <- read_delim("~/Projects/Sediments/MAG_table_further_analysis.tsv")

d2 <- amp_load(otutable = myotutable %>% select(-ends_with("DNA")),
              metadata = mymetadata %>% mutate(sediment_depth_cm = as.numeric(gsub(",",".",sediment_depth_cm)), Bioturbation = ifelse(sediment_depth_cm < 7,"Bioturbated", "Non-bioturbated")))
asv_heatmap <- amp_heatmap(d2,
            group_by = "sediment_depth_cm",
            facet_by = "Bioturbation",
            tax_aggregate = "Genus",
            tax_add = "Phylum",
            tax_show = 20,
            color_vector = c("lightblue", "darkmagenta"),
            plot_colorscale = "sqrt",
            plot_values = FALSE) +
  #theme(text=element_text(family="Helvetica", size=22))+
  theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
        axis.text.y = element_text(size=8),
        legend.position="right")
asv_heatmap
```


```{r}
#mags_further_analysis = mags_further_analysis %>% mutate(Genus = ifelse(is.na(Genus), "NA", Genus)) %>% rowwise()  %>% mutate( taxon = paste(Order,Family,Genus,collapse = "_"))
# top10_surface = lill_tax2 %>% filter(DOMAIN == "Bacteria", Type =="RNA", Depth < 7) %>% group_by(Genus) %>% summarise(nreads = sum(count)) %>% slice_max(nreads, n=10) 
# myotutable2 <- myotutable %>% select(-ends_with("DNA")) %>% filter(Genus %in% mags_further_analysis$Genus)
# top10_bottom = lill_tax2 %>% filter(DOMAIN == "Bacteria", Type =="RNA", Depth < 7) %>% group_by(Genus) %>% summarise(nreads = sum(count)) %>% slice_max(nreads, n=10) 
# myotutable2 <- myotutable %>% select(-ends_with("DNA")) %>% filter(Genus %in% mags_further_analysis$Genus)
# #myotutable2 <-myotutable2 %>% filter(taxon %in% gsub(" ","",mags_further_analysis$taxon))
# d3 <- amp_load(otutable = myotutable2 ,
#               metadata = mymetadata %>% mutate(sediment_depth_cm = as.numeric(gsub(",",".",sediment_depth_cm)), Bioturbation = ifelse(sediment_depth_cm < 7,"Bioturbated", "Non-bioturbated")))
# amp_heatmap(d3,
#             group_by = "sediment_depth_cm",
#             facet_by = "Bioturbation",
#             tax_aggregate = "Genus",
#             tax_add = "Phylum",
#             color_vector = c("blue", "red"),
#             plot_colorscale = "sqrt",
#             plot_values = FALSE) +
#   theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
#         axis.text.y = element_text(size=8),
#         legend.position="right")
```


```{r}
raw_genome_1 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_1_genome_info.tsv",delim="\t") %>% mutate(reads_source = 1)
raw_genome_3 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_3_genome_info.tsv",delim="\t") %>% mutate(reads_source = 3)
raw_genome_7 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_7_genome_info.tsv",delim="\t") %>% mutate(reads_source = 7)
raw_genome_10 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_10_genome_info.tsv",delim="\t") %>% mutate(reads_source = 10)
raw_genome_25 <- read_delim("~/Projects/Sediments/all_depths/IS_raw_M5_3_25_genome_info.tsv",delim="\t") %>% mutate(reads_source = 25)
lifestyle_plus_phylo <- read_delim("~/Projects/Sediments/lifestyles.tsv",delim="\t")
raw_genome_M5_3 <- rbind(raw_genome_1, raw_genome_3, raw_genome_7, raw_genome_10, raw_genome_25)

raw_genome_1 <- read_delim("~/Projects/Sediments/all_depths_M1_3/IS_raw_M1_3_1_genome_info.tsv",delim="\t") %>% mutate(reads_source = 1)
raw_genome_3 <- read_delim("~/Projects/Sediments/all_depths_M1_3/IS_raw_M1_3_3_genome_info.tsv",delim="\t") %>% mutate(reads_source = 3)
raw_genome_7 <- read_delim("~/Projects/Sediments/all_depths_M1_3/IS_raw_M1_3_7_genome_info.tsv",delim="\t") %>% mutate(reads_source = 7)
raw_genome_10 <- read_delim("~/Projects/Sediments/all_depths_M1_3/IS_raw_M1_3_10_genome_info.tsv",delim="\t") %>% mutate(reads_source = 10)
raw_genome_25 <- read_delim("~/Projects/Sediments/all_depths_M1_3/IS_raw_M1_3_25_genome_info.tsv",delim="\t") %>% mutate(reads_source = 25)
lifestyle_plus_phylo_M1 <- read_delim("~/Projects/Sediments/lifestyles_M1_3.tsv",delim="\t")
raw_genome_M1_3 <- rbind(raw_genome_1, raw_genome_3, raw_genome_7, raw_genome_10, raw_genome_25)

left_join(raw_genome_M5_3, lifestyle_plus_phylo)

mags_further_analysis <- mags_further_analysis %>% mutate(genome = gsub(".filtered.fa","",MAG)) %>% filter(!is.na(Family))
mag_ab_plot <- left_join(
  rbind(raw_genome_M5_3,raw_genome_M1_3) %>%
    mutate(genome = gsub(".filtered.fa","",gsub("refs/","",genome))),
  mags_further_analysis) %>%
  mutate(Genus = ifelse((Genus == "NA") | is.na(Genus),paste("Uncl.", Family),Genus)) %>%
  mutate(name = paste0(genome, " (", Genus, ")")) %>%
  mutate(site = gsub("_.*","",genome)) %>%
  ggplot(aes(x=factor(reads_source), y=name, fill=log10(coverage))) +
    geom_tile(width=0.9) +
    scale_fill_continuous(high="darkmagenta",low="lightblue") +
    scale_x_discrete(guide = guide_axis(angle = 60)) +
    theme(axis.title = element_text(size=22), legend.title = element_text(size=16), axis.text.y=element_text(size=14)) +
    labs(y="MAGs", x="Reads Source (CmBSF)", color="Log10 Coverage") +
    scale_y_discrete(limits=rev) +
    facet_wrap(~site,scales="free_y",ncol=1) +
    theme(strip.text.x = element_blank())

mag_ab_plot
```
Here I generate the tree using the ggtree library, I remove MAGs with >15% contamination or <5% completeness from the coloring and abundance 
```{r}
m1_coverages <- rbind(
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_1.txt") %>% mutate(reads_source = 1),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_3.txt") %>% mutate(reads_source = 3),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_5.txt") %>% mutate(reads_source = 5),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_7.txt") %>% mutate(reads_source = 7),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_10.txt") %>% mutate(reads_source = 10),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M1_3_25.txt") %>% mutate(reads_source = 25)
    ) %>% mutate(site = "M1")

m5_coverages <- rbind(
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_1.txt") %>% mutate(reads_source = 1),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_3.txt") %>% mutate(reads_source = 3),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_5.txt") %>% mutate(reads_source = 5),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_7.txt") %>% mutate(reads_source = 7),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_10.txt") %>% mutate(reads_source = 10),
    read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/coverage_M5_3_25.txt") %>% mutate(reads_source = 25)
    ) %>% mutate(site = "M5")

stb_m5_3 <- read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/combined_reference_M5_3.stb",col_names = c("scaffold","genome"))
stb_m1_3 <- read_delim("~/Projects/Sediments/all_mags_all_depths/coverages/combined_reference_M1_3.stb",col_names = c("scaffold","genome"))
new_abundances <- rbind(
  left_join(m1_coverages %>% rename(scaffold = `#rname`), stb_m1_3),
  left_join(m5_coverages %>% rename(scaffold = `#rname`), stb_m5_3)
  )
new_coverages <- new_abundances %>%
  select(genome, scaffold, site, reads_source, meandepth, endpos) %>%
  rename(scaff_len = endpos) %>%
  pivot_wider(
    id_cols=c(genome, site, scaffold, scaff_len),
    names_from = reads_source,
    values_from = meandepth,
    values_fill = 0,
    names_prefix = "cov_depth_") %>% 
  pivot_longer(cols = starts_with("cov_depth_"), names_to = "reads_source", values_to = "meandepth", names_prefix = "cov_depth_") %>% 
  mutate(reads_source = as.integer(reads_source)) %>%
  group_by(genome, site, reads_source) %>% 
  summarise(
    mean_cov = weighted.mean(meandepth, scaff_len),
    length = sum(scaff_len)
  ) 

```
```{r}
new_coverages %>% filter((genome %in% raw_genome_M1_3$genome)|(paste0("refs/",genome) %in% raw_genome_M5_3$genome)) %>% ggplot(aes(x=as.factor(reads_source), y = genome, fill=mean_cov)) + geom_tile() + colorspace::scale_fill_continuous_sequential(palette = "ag_Sunset") + facet_wrap(~site, scales="free")
```

```{r}
summed_len = sum(new_coverages$length)
summed_covg = sum(new_coverages$mean_cov)
new_rel_abs = new_coverages %>% mutate(genome = str_extract(genome,"M[15]_3.[0-9]+")) %>% mutate(ra = mean_cov / summed_covg) %>% pivot_wider(id_cols=c(genome, site), names_from = reads_source, values_from = c(ra, mean_cov))
```
```{r}
linnean = c("domain","phylum","class","order","family","genus","species")
phylogeny_M5_3 <- read_delim("~/Projects/Sediments/M5_3.gtdbtk.bac120.summary.tsv")
phylogeny_M1_3 <- read_delim("~/Projects/Sediments/M1_3.gtdbtk.bac120.summary.tsv")
phylogeny <- rbind(phylogeny_M1_3, phylogeny_M5_3) %>%
  mutate(genome = str_extract(user_genome,"M[15]_3.[0-9]+")) %>%
  select(genome, classification) %>%
  separate_wider_delim(
    cols=classification,
    delim=";",
    names =linnean,
    too_few="align_start"
    ) %>%
  mutate(across( all_of(linnean), ~str_extract(.x,"(?<=[a-z]__).+"))) %>% 
  mutate(phylum = ifelse(is.na(phylum),"Unclassified",phylum))

new_rel_abs <- left_join(new_rel_abs, phylogeny) %>% ungroup()
```
```{r}
quality_m5_3 <- read_delim("~/Projects/Sediments/checkm2_results_M5_3_all.tsv")
quality_m1_3 <- read_delim("~/Projects/Sediments/checkm2_results_M1_3_all.tsv")
quality <- rbind(quality_m5_3, quality_m1_3) %>% mutate(genome =str_extract(Name,"M[15]_3.[0-9]+") ) %>% select(genome, Completeness, Contamination)

new_rel_abs <- left_join(new_rel_abs, quality)
```



```{r}
library("treeio")
library("ggtree")

levels_ <- c("Acidobacteriota","Actinobacteriota","Atribacterota","Bacteroidota","Bipolaricaulota","BMS3Abin14","Calditrichota","Campylobacterota","Chloroflexota","Desulfobacterota","Gemmatimonadota","Hydrogenedentota","JABDJQ01","Krumholzibacteriota","KSB1","Methylomirabilota","Myxococcota","Myxococcota_A","Nitrospirota","Patescibacteria","Planctomycetota","Poribacteria","Proteobacteria","Spirochaetota","Verrucomicrobiota","Zixibacteria")
polychrome_27 <- c("#358d78","#a3583b","#b400fe","#fb0000","#00fe16","#fe1cbe","#f9bbfb","#ece500","#2aabff","#ff781c","#eedfb2","#45427e","#1cfdb4","#c1004f","#6aa526","#841662","#1ce8fd","#fc72fe","#faaf00","#322ac4","#937881","#cae5f6","#870da0","#fd90b4","#685d0d","#aff822","#b19afc")
names(polychrome_27) <- levels_

#m5_3_quality <- read_delim("~/Projects/Sediments/M5_3_MAG_completeness_contamination.tsv",delim="\t") %>% mutate(user_genome = gsub(".fa","",MAG))
#m5_25_quality <- read_delim("~/Projects/Sediments/M5_25_MAG_completeness_contamination.tsv",delim="\t") %>% mutate(user_genome = gsub(".fa","",MAG))

sed_tree <- read.tree("~/Projects/Sediments/combined_sites_bacteria_tree_vft.nwk")
sed_tree$tip.label = gsub(".filtered","",sed_tree$tip.label)
# m5_abundances <- read.table("~/Projects/Sediments/combined_M5_w_abundance.tsv",header=T, sep="\t")
# m5_abundances <- left_join(m5_abundances, rbind(m5_3_quality,m5_25_quality))
# m5_abundances_f <- m5_abundances %>% filter(Duplications < 15, Completeness > 5)
# m5_abundances_f <- m5_abundances_f %>%
#   mutate(
#     Phylum = gsub("p__","",Phylum),
#     Class = gsub("c__","",Class),
#     Family = gsub("f__","",Family),
#     Genus =gsub("g__","",Genus),
#     Species = gsub("s__","",Species)
#     ) %>% mutate(user_genome = gsub(".filtered","",user_genome))
temp2 <- new_rel_abs %>% group_split(phylum)
l5 <- list()
for (tib in temp2){
    l5[[tib$phylum[1]]] = tib$genome
}

sed_tree_w_phyla <- groupOTU(sed_tree, l5, "Phylum_")
sed_tree_colored <- ggtree(sed_tree_w_phyla, layout="circular",aes(color=Phylum_),key_glyph="rect")
new_rel_abs2 <- new_rel_abs %>% select("genome",starts_with("ra")) %>% column_to_rownames("genome")
sed_tree_w_ab <- gheatmap(sed_tree_colored, data=new_rel_abs2, high="darkmagenta",low="lightblue",width = 0.1, font.size = 0.1) 
# ctb <- read_table("~/Projects/Sediments/all_depths/combined_reference_M5_3.stb",col_names = c("scaffold","MAG"))
# ctb <- ctb %>% mutate(MAG = str_replace(MAG, ".filtered.fa",""))
sed_tree_w_ab_circle <- sed_tree_w_ab + geom_tiplab(aes(subset=(label %in% c(mags_further_analysis$genome))),offset=0.17, size=4, align=T, linesize = 0.01) + scale_color_manual(name="Phylum",values=polychrome_27) + theme(legend.position="right", legend.spacing.y = unit(0,"mm")) + guides(
    size = guide_legend(
      override.aes = list(linetype = c("solid", "blank"), shape = c(NA, 16))
    )
  ) 
sed_tree_w_ab_circle
```

Here I use cowplot to join all the figures together into a panel
```{r}
sediment_cartoon <- ggplot() + cowplot::draw_image("~/Documents/Sediments presentations + paper/Figures/Sediment_core_cartoon_v2.svg") + cowplot::panel_border(remove=T)
#legend = cowplot::get_plot_component(m5_combined_tree_w_ab_circle, 'guide-box-bottom', return_all = TRUE)
#rna_plus_legend <- cowplot::plot_grid(rna_16s_plot, legend, ncol=1, nrow=2, rel_heights = c(1.5, 0.5))
bottom_row = cowplot::plot_grid(sed_tree_w_ab_circle + theme(legend.position="none") , cowplot::get_legend(sed_tree_w_ab_circle), mag_ab_plot, rel_widths = c(1,0.1,1),rel_heights = c(0.9, 0.5, 1), label_size = 32,labels = c("D","","E"), nrow=1) 
top_row = cowplot::plot_grid(sediment_cartoon, nmds_plot, asv_heatmap, label_size = 32,nrow=1, rel_widths = c(0.8,0.7,1),labels=c("A","B","C"))
panel_1 <- cowplot::plot_grid(top_row, bottom_row, nrow = 2, ncol=1)
panel_1
#ggplot2::ggsave("~/Documents/Sediments presentations + paper/Figures/panel_1_draft_2.svg", panel_1,width =2300,height=1300, units="px")# these render funky so I use the normal Rstudio save instead
#ggplot2::ggsave("~/Documents/Sediments presentations + paper/Figures/panel_1_draft_2.png", panel_1,width =2300,height=1300, units="px")
```
```{r}
rna_otu_counts <- myotutable %>%
  select(OTU, ends_with("RNA")) %>%
  pivot_longer(
    cols = ends_with("RNA"),
    names_to = "read_source",
    values_to = "read_count",
    names_pattern = "(?<=_)([0-9]+x?5?)(?=_RNA)"
    ) %>%
  rowwise() %>%
  mutate(read_source = as.numeric(gsub("x",".",read_source)))

rna_otus_bt_vs_nbt <- rna_otu_counts %>% mutate(bioturb = ifelse(read_source <= 6, "bt","nbt")) %>% group_by(OTU,bioturb) %>% summarise(sum_reads = sum(read_count)) %>% ungroup()
rna_otus_bt_vs_nbt_w <- rna_otus_bt_vs_nbt %>% pivot_wider(id_cols = OTU, names_from = "bioturb", values_from = "sum_reads")
```
```{r}
rna_otus_bt_vs_nbt_w %>% mutate(present_bt = bt >= 10, present_nbt = nbt >= 10) %>% group_by(present_bt, present_nbt) %>% summarise(n = n())
```

