---
title: "selection_by_genes"
output: html_document
date: "2025-04-07"
---

```{r}
library(tidyverse)
library(colorspace)
```

```{r}
rgwt_m5 <- read_delim("~/Projects/Sediments/raw_genes_filtered_w_theta_hat_M5.tsv") %>% mutate(PnPs = ifelse(SNV_count > 2, P_n / P_s, NA))  %>% mutate(PnPs = ifelse(is.infinite(PnPs),NA,PnPs))
head(rgwt_m5)
```

```{r}
rgwt_m5 %>% ggplot(aes(x=S_SNV_Site_count)) + geom_histogram(binwidth = 1) + xlim(0,100) 
```

This actually follows a negative binomial with 1 "success" per gene and prob = mean # of Syn SNV sites (21.78) / mean Syn. length (216.5nt)  well .  Could maybe do with some fine tuning but I'm not sure what I'd use that info for atm :)
```{r}
ggplot() + geom_histogram(aes(x=rnbinom(n=50000, 1, prob=0.1)), binwidth = 1) + xlim(0,100)
```


```{r}
rgwt_m1 <- read_delim("~/Projects/Sediments/raw_genes_filtered_w_theta_hat_m1_3.tsv")%>% mutate(PnPs = ifelse(SNV_count > 2, P_n / P_s, NA)) %>% mutate(PnPs = ifelse(is.infinite(PnPs),NA,PnPs))

```
```{r}
lifestyle_plus_phylo_m5 <- read_delim("~/Projects/Sediments/lifestyles.tsv")
head(lifestyle_plus_phylo_m5)
```

```{r}
lifestyle_plus_phylo_m1 <- read_delim("~/Projects/Sediments/lifestyles_M1_3.tsv")
head(lifestyle_plus_phylo_m1)
```

```{r}
rgwt_m5 <- left_join(rgwt_m5 %>% rowwise() %>% mutate(genome = str_extract(MAG,"M[15]_[0-9].[0-9]+")) %>% ungroup(),lifestyle_plus_phylo_m5)
rgwt_m1 <- left_join(rgwt_m1 %>% rowwise() %>% mutate(genome = str_extract(MAG,"M[15]_[0-9].[0-9]+")) %>% ungroup(),lifestyle_plus_phylo_m1)
```

```{r}
combined_genes_3_25 <- rbind(rgwt_m5 %>% mutate(site="M5"), rgwt_m1 %>% mutate(site="M1")) %>%
  ungroup() %>%
  filter((coverage >= 5) & ((reads_source == 3) | (reads_source == 25))) %>%
  filter(!is.na(lifestyle)) %>%
  select(gene, genome, lifestyle, site, PnPs, reads_source) %>%
  pivot_wider(names_from = reads_source, names_prefix = "PnPs_", values_from = PnPs, id_cols=c(gene, genome,site, lifestyle), values_fill = NA)
```


```{r}
unique(rgwt_m5$lifestyle)
```

the palette names come from colorspace sequential sets (I'm loosely following the red orange purple scheme found in Sunset but splitting it by lifestyle)

```{r}
generate_hex_plot <- function(plot_site, plot_lifestyle, plot_palette, plot_title, plot_df){
  ph = sequential_hcl(2, palette=plot_palette)[1] # grab the high and low colors from the colorspace library
  pl = sequential_hcl(2, palette=plot_palette)[2]
  return(#there is 0 need to save this object inside the function
    plot_df %>%
      filter((lifestyle == plot_lifestyle) & (site == plot_site)) %>%
      ggplot(aes(x=PnPs_25, y=PnPs_3)) +
        geom_hex() +
        labs(x="πN/πS 25 CmBSF", y="πN/πS 3 CmBSF",title=plot_title) +
        xlim(0,1) +
        ylim(0,1) +
        coord_fixed() + 
        geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
        scale_fill_continuous(high=ph,low=pl) # setting the color with aes fill would set each hex the same and remove density info
  )
}
 
```

```{r}
m5_sp <- generate_hex_plot("M5", "Surface Specialist","RedOr","M5 Surface Spec.",combined_genes_3_25)
```
```{r}
m5_sp
```

```{r}
m5_gen <- generate_hex_plot("M5", "Generalist","PurpOr","Generalist",combined_genes_3_25)
m5_ssp <- generate_hex_plot("M5", "Subsurface Specialist","Purp","Subsurf. Spec.",combined_genes_3_25)
m1_sp <- generate_hex_plot("M1", "Surface Specialist","RedOr","M1",combined_genes_3_25)
m1_gen <- generate_hex_plot("M1", "Generalist","PurpOr","",combined_genes_3_25)
m1_ssp <- generate_hex_plot("M1", "Subsurface Specialist","Purp","",combined_genes_3_25)
```

```{r}
m5_row <- cowplot::plot_grid(m5_sp, m5_gen, m5_ssp, nrow=1)
m1_row <- cowplot::plot_grid(m1_sp, m1_gen, m1_ssp, nrow=1)
cowplot::plot_grid(m5_row,m1_row,ncol=1)
```
```{r}
busco_gene_ids_M5 <- read_delim("~/Projects/Sediments/busco_gene_ids.txt",delim=",")
head(busco_gene_ids_M5)
```

```{r}
busco_gene_ids_M1 <- read_delim("~/Projects/Sediments/M1_3_busco_gene_ids.txt",delim=",",col_names = c("gene"))
head(busco_gene_ids_M1)
```
```{r}
rgwt_m5 <- rgwt_m5 %>% mutate(is_busco = gene %in% busco_gene_ids_M5$gene) 
rgwt_m1 <- rgwt_m1 %>% mutate(is_busco = gene %in% busco_gene_ids_M1$gene)
```

```{r}
sequential_hcl(2, palette="ag_Sunset") 
```

```{r}
rgwt_m5 %>% 
  filter((coverage >= 5) & ((reads_source == 3) | (reads_source == 25))) %>%
  select(gene, genome, is_busco, PnPs, reads_source) %>%
  pivot_wider(
    names_from = reads_source,
    names_prefix = "PnPs_",
    values_from = PnPs,
    id_cols=c(gene, genome, is_busco),
    values_fill = NA) %>%
  ggplot(aes(x=PnPs_25, y=PnPs_3)) +
    geom_hex() +
    labs(x="πN/πS 25 CmBSF", y="πN/πS 3 CmBSF") +
    xlim(0,1) +
    ylim(0,1) +
    coord_fixed() + 
    geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
    scale_fill_continuous(high="#4B1D91",low="#E7D39A") +
    facet_grid(is_busco ~ genome)
  
  
```
```{r}
rgwt_m1 %>% 
  filter((coverage >= 5) & ((reads_source == 3) | (reads_source == 25))) %>%
  select(gene, genome, is_busco, PnPs, reads_source) %>%
  pivot_wider(
    names_from = reads_source,
    names_prefix = "PnPs_",
    values_from = PnPs,
    id_cols=c(gene, genome, is_busco),
    values_fill = NA) %>%
  ggplot(aes(x=PnPs_25, y=PnPs_3)) +
    geom_hex() +
    labs(x="πN/πS 25 CmBSF", y="πN/πS 3 CmBSF") +
    xlim(0,1) +
    ylim(0,1) +
    coord_fixed() + 
    geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
    scale_fill_continuous(high="#4B1D91",low="#E7D39A") +
    facet_grid(is_busco ~ genome)
```
```{r}
unlogged_pnps_distr <- rgwt_m5 %>%
  ggplot(aes(x=factor(reads_source), y=PnPs, color=is_busco)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  scale_color_discrete_qualitative(palette = "Dark3") + 
  labs(y="πN/πS", x="Cm Below Seafloor") +
  theme(legend.position="none") +
  ylim(0,1) + 
  facet_wrap(~genome)
```

```{r}
logged_pnps_distr <- rgwt_m5 %>% 
  ggplot(aes(x=factor(reads_source), y=log10(PnPs+0.01), color=is_busco)) +
  geom_violin(draw_quantiles = c(0.5)) +
  scale_color_discrete_qualitative(palette = "Dark3") + 
  labs(y="Log10 πN/πS + 10^-2",x="Cm Below Seafloor",color="BUSCO status") + 
  facet_wrap(~genome) +
  theme(legend.position = c(.9, .1), axis.title.x=element_blank())

```

```{r}
cowplot::plot_grid(logged_pnps_distr, unlogged_pnps_distr, ncol=1)
```
```{r}
 bmcs <-  rgwt_m5 %>% filter(is_busco) %>% group_by(MAG, reads_source) %>% summarise(bmc = mean(coverage, na.rm=T)) %>% ungroup()

gene_corr_coefs <- left_join(
  rgwt_m5 %>%
    pivot_wider(
      id_cols=c("MAG","gene"),
      names_from=reads_source,
      values_from=c(coverage),
      names_prefix = "coverage_") %>%
    pivot_longer(
      cols=starts_with("coverage_"),
      names_to="reads_source",
      values_to = "coverage") %>%
    mutate(
      coverage = ifelse(is.na(coverage),0,coverage),
      reads_source=as.integer(gsub("coverage_","",reads_source))),
  bmcs) %>%
  group_by(MAG, gene) %>%
  #slice_head(n=10)
  summarise(busco_corr = cor(coverage, bmc), covg_var = var(coverage / bmc))

gene_corr_coefs <- gene_corr_coefs %>% mutate(is_busco = gene %in% busco_gene_ids_M5$gene)
head(gene_corr_coefs)
```

```{r}
gene_corr_coefs %>% filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>% ggplot(aes(x=busco_corr)) + geom_histogram(binwidth=0.05) + facet_wrap(~MAG) + labs(x="Pearsons r with BUSCO genes") + geom_vline(xintercept=0.9, linetype="dotted")
```
```{r}
gene_corr_coefs %>% mutate(busco_status = ifelse(is_busco, "BUSCO","Others")) %>% filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>% ggplot(aes(x=busco_corr)) + geom_histogram(binwidth=0.05) + facet_wrap(is_busco~MAG) + labs(x="Pearsons r with BUSCO genes") + geom_vline(xintercept=0.75, linetype="dotted") + facet_grid(MAG~busco_status,scale="free_y")
```
```{r}
gene_corr_coefs %>% mutate(busco_status = ifelse(is_busco, "BUSCO","Others")) %>% filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>% ggplot(aes(x=covg_var)) + geom_histogram(binwidth=0.05) + facet_grid(MAG ~busco_status,scales="free_y") + xlim(0,1)
```


looking at it another way
```{r}
rgwt_m5 %>%
    pivot_wider(
      id_cols=c("MAG","gene","is_busco"),
      names_from=reads_source,
      values_from=c(coverage),
      names_prefix = "coverage_") %>%
    pivot_longer(
      cols=starts_with("coverage_"),
      names_to="reads_source",
      values_to = "coverage") %>%
    mutate(
      coverage = ifelse(is.na(coverage),0,coverage),
      reads_source=as.integer(gsub("coverage_","",reads_source))) %>%
  filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
  mutate(busco_status = ifelse(is_busco, "BUSCO","Others")) %>%
  ggplot(aes(x=busco_status, y=log10(coverage+0.01), color=is_busco)) +
    geom_violin(draw_quantiles = c(0.5)) +
    facet_wrap(~MAG) +
    scale_color_discrete_qualitative(palette = "Dark3") + 
    labs(x="BUSCO status", y="log10 Coverage + 1E-2") +
    theme(legend.position = "None")
```
sweet some more poisson/negative binomial like stats :)

```{r}
gene_corr_coefs %>% filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>% group_by(MAG, is_busco) %>% summarise(puttv_core_ct = sum(busco_corr >= 0.75), puttv_accessory_ct = sum(busco_corr < 0.75))
```
```{r}
df = data.frame(MAG=c(),puttv_core_ct_FALSE=c(),puttv_core_ct_TRUE=c(), puttv_accessory_ct_FALSE = c(), puttv_accessory_ct_TRUE = c(), n_busco = c(), fraction_BUSCO_core = c(), fraction_BUSCO_acc =c(), fraction_others_core = c(), fraction_others_acc =c(), cutoff=c())
for(cutoff in seq(0, 1, 0.05)){
  df = rbind(df , gene_corr_coefs %>%
    filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
    group_by(MAG, is_busco) %>% 
    summarise(puttv_core_ct = sum(busco_corr >= cutoff), puttv_accessory_ct = sum(busco_corr < cutoff)) %>%
    ungroup() %>%
    pivot_wider(names_from = is_busco, values_from=c(puttv_core_ct, puttv_accessory_ct), id_cols=c(MAG)) %>% 
    mutate(
      n_busco = puttv_core_ct_TRUE + puttv_accessory_ct_TRUE,
      fraction_BUSCO_core = (puttv_core_ct_TRUE / (puttv_core_ct_TRUE + puttv_accessory_ct_TRUE)), 
      fraction_BUSCO_acc = (puttv_accessory_ct_TRUE / (puttv_core_ct_TRUE + puttv_accessory_ct_TRUE)), 
      fraction_others_core = (puttv_core_ct_FALSE / (puttv_core_ct_FALSE + puttv_accessory_ct_FALSE)), 
      fraction_others_acc = (puttv_accessory_ct_FALSE / (puttv_core_ct_FALSE + puttv_accessory_ct_FALSE)),
      cutoff = cutoff
    ))
}  
```
```{r}
df %>% ggplot(aes(x=cutoff, y=fraction_BUSCO_core, color=MAG)) + geom_point() + scale_color_discrete_sequential(palette="ag_Sunset") + labs(y="Fraction BUSCO assigned to Core",x="r cutoff")
```

```{r}
rgwt_m5_wcorr <- left_join(rgwt_m5, gene_corr_coefs) %>% mutate(status = ifelse((busco_corr < 0.75) | (is_busco),"Acc","Core"))
```

```{r}
rgwt_m5_wcorr %>% 
  filter((coverage >= 5) & ((reads_source == 3) | (reads_source == 25))) %>%
  filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
  select(gene, genome, is_busco, PnPs, reads_source, status) %>%
  pivot_wider(
    names_from = reads_source,
    names_prefix = "PnPs_",
    values_from = PnPs,
    id_cols=c(gene, genome, is_busco, status),
    values_fill = NA) %>%
  ggplot(aes(x=PnPs_25, y=PnPs_3)) +
    geom_hex() +
    labs(x="πN/πS 25 CmBSF", y="πN/πS 3 CmBSF") +
    xlim(0,1) +
    ylim(0,1) +
    coord_fixed() + 
    geom_abline(slope=1, intercept=0, color="black",linewidth=1, linetype="dotted") +
    scale_fill_continuous(high="#4B1D91",low="#E7D39A") +
    facet_grid(status ~ genome)
  
```

```{r}
bootstrap_gene <- function(snv_s_count, snv_n_count, ln, ls, covg){
  nrep_ = 1000
  pns = rep(NA, nrep_)
  for(i in seq(1:(1+nrep_))){
    s = rpois(1, snv_s_count)
    n = rpois(1, snv_n_count)
    p_n = n / (ln * log(covg))
    p_s = s / (ls * log(covg))
    pnps = p_n / p_s
    pns[i] = pnps
  }
  return(c(mean(pns), sd(pns)))
}
bootstrap_gene_v <- Vectorize(bootstrap_gene)
```
```{r}
output <- rgwt_m5_wcorr %>% filter((PnPs >= 0.9) & (PnPs <= 1.1)) %>% rowwise() %>% mutate(bs_mean = bootstrap_gene(SNV_S_count, SNV_N_count, Ln, Ls, coverage)[1], bs_std = bootstrap_gene(SNV_S_count, SNV_N_count, Ln, Ls, coverage)[2]) %>% ungroup()
```
```{r}
output %>% ungroup() %>% filter(!is.infinite(bs_mean)) %>% summarise(mean_of_means = mean(bs_mean, na.rm=T), sd(bs_mean,na.rm=T))
```

```{r}
1.12 - 3*0.085 
```
```{r}
1.12 + 3*0.085
```


```{r}
data.frame(output) %>% ggplot(aes(x=bs_mean)) + geom_histogram()
```


```{r}
core_acc_evol_table <- rgwt_m5_wcorr %>%
  filter(MAG %in% c("M5_3.3.filtered.fa","M5_3.252.filtered.fa","M5_3.68.filtered.fa")) %>%
  rename(pang_status = status) %>% #pangenome status
  mutate(evol_status = ifelse(PnPs < 0.865, "Conserved", ifelse(PnPs < 1.375, "Neutral", "Positive"))) %>%
  group_by(genome, pang_status, evol_status, reads_source) %>%
  summarise(n = n()) %>% 
  ungroup() 
core_acc_evol_table %>%
  pivot_wider(values_from=n, names_from = c(pang_status, evol_status))
```

```{r}
core_acc_evol_table
```

```{r}
#core_acc_evol_table %>%filter(!is.na(evol_status) & !is.na(pang_status))  %>% uncount(n) %>% group_by(reads_source, genome) %>% summarise(p_val = chisq.test(pang_status, evol_status,simulate.p.value = T)$p.value) %>% pivot_wider(values_from=p_val, names_from = reads_source, names_prefix="sample_")
```


```{r}
core_acc_evol_table_w <- core_acc_evol_table %>% pivot_wider(values_from=n, names_from = c(pang_status, evol_status))
```

```{r}
core_acc_evol_table_w
```
```{r}
options(digits=3)
core_acc_evol_table_w %>%
  ungroup() %>%
  select(-ends_with("NA")) %>% 
  mutate(ngenes = rowSums(across(-c(genome,reads_source)))) %>%
  mutate(across(-c(genome,reads_source, ngenes))/rowSums(across(-c(genome,reads_source, ngenes))))
```

```{r}
options(digits=3)
core_acc_evol_table_w %>%
  ungroup() %>%
  select(-ends_with("NA")) %>% 
  mutate(ngenes = rowSums(across(-c(genome,reads_source)))) %>%
  mutate(across(starts_with("Acc"))/rowSums(across(starts_with("Acc")))) %>%
  mutate(across(starts_with("Core"))/rowSums(across(starts_with("Core")))) 
```


```{r}
chisq.test(data.matrix(core_acc_evol_table_w %>% select(-ends_with("NA")) %>% select(-reads_source) %>% filter(genome == "M5_3.252")),simulate.p.value = T)
```
```{r}
chisq.test(data.matrix(core_acc_evol_table_w %>% select(-ends_with("NA")) %>% select(-reads_source) %>% filter(genome == "M5_3.3")),simulate.p.value = T)
```
```{r}
chisq.test(data.matrix(core_acc_evol_table_w %>% select(-ends_with("NA")) %>% filter(reads_source != 25) %>% select(-reads_source) %>% filter(genome == "M5_3.68")),simulate.p.value = T)
```
```{r}
chisq.test(data.matrix(core_acc_evol_table_w %>% select(-ends_with("NA"))  %>% select(-reads_source) %>% filter(genome == "M5_3.68")),simulate.p.value = T)
```

```{r}
core_acc_evol_table %>% filter(genome == "M5_3.68") %>%  pivot_wider(values_from=n, names_from = c(pang_status, evol_status)) %>% select(-ends_with("NA"))
```


```{r}
core_acc_evol_table_w %>% write_delim("~/Projects/Sediments/genes_by_evol_and_pang_status.tsv",delim="\t")
```

```{r}
proc_m5_3 <- read_delim("~/Projects/Sediments/proc_genes_filtered_w_theta_hat_M5.tsv")
proc_m5_3 <- left_join(proc_m5_3 %>% rowwise() %>% mutate(genome = str_extract(MAG,"M[15]_[0-9].[0-9]+")) %>% ungroup(),lifestyle_plus_phylo_m5)
proc_m5_3 <- proc_m5_3  %>% mutate(is_busco = gene %in% busco_gene_ids_M5$gene) 
```

```{r}
logged_pnps_distr_proc <- proc_m5_3 %>% 
  
  ggplot(aes(x=factor(reads_source), y=log10(pn_ps+0.01), color=is_busco)) +
  geom_violin(draw_quantiles = c(0.5)) +
  scale_color_discrete_qualitative(palette = "Dark3") + 
  labs(y="Log10 πN/πS + 10^-2",x="Cm Below Seafloor",color="BUSCO status") + 
  facet_wrap(~genome) +
  theme(legend.position = c(.9, .1), axis.title.x=element_blank())
logged_pnps_distr_proc
```
```{r}
proc_genes_3_25 <- proc_m5_3 %>%
  ungroup() %>%
  filter((coverage >= 5) & ((reads_source == 3) | (reads_source == 25))) %>%
  filter(!is.na(lifestyle)) %>%
  select(gene, genome, lifestyle, pn_ps, reads_source) %>%
  pivot_wider(names_from = reads_source, names_prefix = "PnPs_", values_from = pn_ps, id_cols=c(gene, genome, lifestyle), values_fill = NA) %>%
  mutate(site = "M5")
  
```
```{r}
m5_gen_p <- generate_hex_plot("M5", "Generalist","PurpOr","Generalist",proc_genes_3_25)
m5_ssp_p <- generate_hex_plot("M5", "Subsurface Specialist","Purp","Subsurf. Spec.",proc_genes_3_25)
m5_sp_p <- generate_hex_plot("M5", "Surface Specialist","RedOr","Norm'ed Surf Spec.",proc_genes_3_25)
```

```{r}
m5_row <- cowplot::plot_grid(m5_sp_p, m5_gen_p, m5_ssp_p, nrow=1)
m5_row

```
